% !TEX root = ./arock_pkg_main.tex
\section{Future Work}

New features are still being added to \pkg. The following are some of the features we are exploring.

\subsection{Stochastic algorithms} 

Stochastic algorithms exploit summative structures in problems to produce cheap updates. Our current toolbox does not currently support stochastic algorithms, but can be modified to do so. Such a modification would require stochastic operators. A developement branch will appear on our github.  

\subsection{Cluster computing}
 
 Currently, agents are realized as threads.
 This limits our toolbox to the multicore regime.
 Future releases intend to use MPI to bring \pkg~to cluster computing. The new functionality will be provided in the Driver and the Kernel Layer. 

\subsection{User interface}
 \pkg~requires the user to either use our prepackaged executables, or code moderately in C++.
 A graphic user interace is in development for those who wish avoid interacting directly with C++.
 This will limit the user to built-in functionality.
 In addition, interfaces to Matlab and Python will be provided for algorithms of particular interest.
 
\subsection{Automatic parameter choice}
\pkg~requires the user to choose stepsizes, and number of threads.
In future releases we intend to provide automatic stepsize heuristics. 
Optimizing thread number is a function of current processor usage and computing architecture. 
We intend to provide functions to survey the current architecture and suggest proper levels of parallelism and asynchrony.

\subsection{Block coordinate update}
  Currently \pkg~does not support block coordinate updates (updates consisting of a set of coordinates). Block coordinate updates present a tradeoff between iteration complexity and communication efficiency.
  Automatic block size deduction and block composition (the coordinates forming a block) is an open question.
We plan to explore several heuristics.

\subsection{New fields}

Splitting methods have been used in many fields (see \citep{roland2016some} for a more in depth discussion).
Our current release focuses on optimization, but provides a strong foundation for branching into other fields.
Fruitful fields to explore include:

\begin{itemize}
\item Numerical simulations;
\item Large scale numerical linear algebra;
\item Time varying systems such as initial value problems and partial differential equations.
\end{itemize}


\section{Conclusion }
We have developed \pkg, an easy-to-use open source toolbox for large scale optimization problems.
The toolbox implemented both sequential and parallel algorithms based on operator splitting methods, stochastic methods,
and coordinate update methods. \pkg~is separated into several layers and mimics how a scientist writes down an optimization algorithm. Therefore, it is easy for one to obtain a new algorithm by modifying just one of the layers such as adding a new operator.


New features and applications will be added to \pkg~based upon new research and community input. The software and user guide \repo~will be kept up-to-date and supported.

\section{Acknowledgements}
The work is supported in part by NSF grants ECCS-1462398.