% !TEX root = ./arock_pkg_main.tex
\subsection{Portfolio Optimization}
Assume that we have one unit of capital and $m$ assets to invest on. The $i$th asset has an expected return rate $\xi_i\ge 0$\cut{ for $i\in[m]$}. Our goal is to find a portfolio with the minimal risk such that the expected return is no less than $c$. This problem can be formulated as

\begin{equation*}
\begin{array}{l}
\displaystyle
\Min_x ~ \frac{1}{2} x^\top Q x, \\
\displaystyle
\text{subject to}~ x\ge0, \sum_{i=1}^m x_i\le 1,\, \sum_{i=1}^m\xi_i x_i\ge c,
\end{array}
\end{equation*}
where the objective function is a measure of risk, and the last constraint imposes that the expected return is at least $c$. We apply the coordinate update scheme as shown in Section 5.3.1 of~\citep{PengWuXuYanYin2016_coordinate} to solve it.
We tested two synthetic  problem instances: one has 5,000 assets and the other instance has 30,000 assets.
The vector of expected return rate was sampled from $N(0.01, 1)$ normal distribution.
The matrix $Q$ was set to $\frac{1}{2} (R + R^T) + \sigma \cdot I$, where the entries of $R$ was sampled independently from $N(0, 0.1)$ normal distribution, and $\sigma$ was chosen such that $Q$ was positive definite.
We tested both sync-parallel and async-parallel methods with 100 epochs. They reached similar final objective. We report the timing and speedup results in Table~\ref{tab:port_opt}. One can observe that  MOTAC scales well for both sync-parallel method and async-parallel method. The nice scaling performance of sync-parallel method is due to almost perfect load balancing across the threads and the homogeneous computing environment.
\begin{table}[!h]
\centering
\begin{tabular}{rrrrrr}
\toprule
\multirow{2}{*}{} & &  \multicolumn{2}{c}{async-parallel} & \multicolumn{2}{c}{sync-parallel} \\
\midrule
 & & time (s) & speedup & time (s) & speedup \\
 \midrule
 \multirow{3}{*}{5K assets} & 1 thread~ & 13.95& 1.00 & 13.98& 1.00 \\
  & 4 threads & 3.53& 3.95 &3.54 & 3.95\\
 & 16 threads & 0.99 &14.09 &0.92 & 15.19\\
\midrule
 \multirow{3}{*}{30K assets} & 1 thread~ & 483.20& 1.00& 479.22& 1.00 \\
  & 4 threads & 124.94& 3.86 &123.74 & 3.87\\
 & 16 threads & 31.96 &15.12 &33.11 & 14.47\\
\bottomrule
\end{tabular}
 \caption{\label{tab:port_opt}Results for portfolio optimization. }
\end{table}

